---
title: "Data_Exploration"
author: "Zamberger Bernd"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    df_print: paged
    fig_height: 7
    fig_width: 10
  pdf_document:
    toc: true
    toc_depth: '3'
---
# Preperation

To load the necessary data, the essential parts of the 01_extract_from_R.rmd script will be integrated in this .rmd. 

To define the working directory in an rmarkdwon code, the setwd() function will be evaluated only for the current chunk. To change it for the whole document, the knitr::opts_knit$set() function provides this functionality. 

Source: [bookdown.org, 07.04.2024](https://bookdown.org/yihui/rmarkdown-cookbook/working-directory.html)

```{r preperation}
#clear all data frames from environment
rm(list=ls(all=TRUE)[sapply(mget(ls(all=TRUE)), class) == "data.frame"])
# set directory-paths
# getwd()
home_dir<-"C:/Users/bernd/Dropbox/Masterarbeit_2023/1_Statistical_Learning_Spangl/"
knitr::opts_knit$set(root.dir = home_dir)
```


```{r setup1}
code_dir<- "../code/"
data_dir<-"../data/extracted_from_rdata/"
fig_dir<-"../figures/"
# load data
load("C:/Users/bernd/Dropbox/Masterarbeit_2023/1_Statistical_Learning_Spangl/code/Code_Bernhard/dataPSsub_new3.RData")
# clean up 
keep<-c("home_dir", # directory structure
		"data_dir", # directory structure
		"code_dir", # directory structure
		"dps1200", # data in R
		"dpsDeriv1200" # data in R
		)
rm(list = setdiff(ls(), keep))

```


```{r libraries}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
library(dplyr)
```

This code is developed to visualize all the spectra in an appropriate manner. For this purpose we use the dps1200 dataframe. 
```{r data_definition}

# Save the data as a new dataframe
# data<-dps1200

data <- dpsDeriv1200

head(data)

```

```{r clean-up}

# Correct column names by removing 'X' at the beginning
names(data) <- gsub("X", "", names(data))

head(data)
```
## Plotting of the spectral data

To visualize the data, different plots will be generated: 

* plot the absorbance against the wavelength

```{r plot-spectra}
# save the spectral data as a new dataframe, without the descriptive coloumns
data_plot<-data[-(1:4)]

# Gathering the data 

# data_plot_long <- gather(data_plot, key = "wavelength", value = "intensity") # gather was reworked to pivot_longer
	# gather() is a function. reworked into pivot_longer()

data_plot_long <- pivot_longer(data_plot,cols = everything(),  names_to="Wavelength [nm]", values_to = "Intensity")

# change the datatype to numeric, for the wavelengths to be displayed correctly
data_plot_long$`Wavelength [nm]`<-(as.numeric(data_plot_long$`Wavelength [nm]`))

# for
str(data_plot_long)

# generate the plots for the spectras
ggplot(data_plot_long, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(alpha = 0.005, size = 1) + # , colour = 3) +
  labs(title = "Overview of the spectra",
       x = "Wavelength [nm]",
       y = "Intensity [1]")

# alpha = 0.005 defines the transparency, points overlapping can be better differentiated

ggplot(data_plot_long, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(alpha = 0.01, size = 0.1) + # , colour = 3) +
  labs(title = "Overview of the spectra",
       x = "Wavelength [nm]",
       y = "Intensity [1]")

```

```{r plot-2columns}
ggplot(data, aes(x = tree)) +
 geom_line(aes(y = `2970`, color = "2970")) +
 geom_line(aes(y = `2883`, color = "2883")) +
 scale_color_manual(values = c("2970" = "blue", "2883" = "red")) +
 labs(x = "Year", y = "Value", color = "Column", title = "")+
	theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```
```{r plot-count1}
# Generate a faceted bar plot by Origin
ggplot(data, aes(x = type)) +
 geom_bar() +
 facet_wrap(~Origin) +
 labs(x = "Type", y = "Count", title = "Count of Types by Country")

# Create the faceted bar plot by Type of Environment
ggplot(data, aes(x = Origin)) +
 geom_bar() +
 facet_wrap(~type) +
 labs(x = "Origin", y = "Count", title = "Count of Origins by Type of Environment")
```

```{r plot-Count-by-Country}

# Create the single bar plot colored by country
ggplot(data, aes(x = type, fill = Origin)) +
 geom_bar(position ="dodge2") +
 labs(x = "Type", y = "Count", fill = "Country", title = "Count of Types by Country")

# Create the single bar plot colored by country


ggplot(data, aes(x = type, fill=Origin)) + 
	geom_bar() +
 labs(x = "Type", y = "Count", fill = "Country", title = "Count of Types by Country (stacked)")
```



```{r num_columns}
# Find the position of the 'type' column
type_position <- which(names(data) == "type")
type_position

# Select all columns after the 'type' column
numerical_columns <- names(data)[type_position + 1:ncol(data)]

# Print the first five selected columns
head(numerical_columns)
```


```{r}
# loop over the unique values in data$type
# create a plot (x= numerical_columns, y = value), color by

data_plot2<-data[-(1:3)]

data_plot_long_2 <- data_plot2 %>%
	pivot_longer(!type ,  names_to="Wavelength [nm]", values_to = "Intensity")

# change the datatype to numeric, for the wavelengths to be displayed correctly
data_plot_long_2$`Wavelength [nm]`<-(as.numeric(data_plot_long_2$`Wavelength [nm]`))

# inspect the plotting data
head(data_plot_long_2)

# generate the plots for the spectra
ggplot(data_plot_long_2, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(alpha=1, size = 0.01, aes(color = type))+ 
  labs(title = "Plot of the spectra, coloured by 'type'",
       x = "Wavelength [nm]",
       y = "Intensity [1]")

# generate plots for the individual type

# living
ggplot(data_plot_long_2, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(data= data_plot_long_2 %>% filter(type == "living") ,
  		   alpha=1, size = 0.01, aes(color = type))+ 
  labs(title = "Test of the spectra",
       x = "Wavelength [nm]",
       y = "Intensity [1]")


# not living
ggplot(data_plot_long_2, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(data= data_plot_long_2 %>% filter(type != "living") ,
  		   alpha=0.1, size = 0.01, aes(color = type))+ 
  labs(title = "NIR-Spectra Intensity by Wavelength and Type",
       x = "Wavelength [nm]",
       y = "Intensity [1]")

```
```{r 2wl_against_each_other}
ggplot(data[4:6], aes(x= data$"2970", y=data$"2968"))+geom_point()

data[5:6]

```

```{r coloured_plot}
head(data)

ggplot(data, aes(x = year, y = `2970`)) +
 geom_point(aes(colour=type)) +
 labs(x = "Year", y = "Value", title = "Scatter Plot of 2970 Over Years")
```


## Plotting of the age/year

for each tree we have multiple spectral measurements, where each measurement includes a specific tree ring. To get a sense for the age distribution, the Quartiles will be calculated. 

```{r} 
#, results='hide'}
ggplot(data, aes(x = year)) +
 geom_bar(fill = "steelblue") +
 theme_minimal() +
	scale_y_continuous(breaks=seq(0, 12, 1))
 labs(x = "Year", y = "Count", title = "Count of Measurements per Date of Wood growth")
```


```{r}
ggplot(data, aes(x = year)) +
 geom_density(fill = "steelblue") +
 theme_minimal() +
 labs(x = "Year", y = "Count", title = "Wood Year Distribution")
```


```{r}
ggplot(data, aes(x = Origin)) +
 geom_bar() +
	theme_minimal() +
 labs(x = "Origin", y = "Count", title = "Measurements per country")
```


Variable variance 

```{r}

plot(sapply(data_plot, var))
```


