---
title: "Data_Exploration"
author: "Zamberger Bernd"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    df_print: paged
    fig_height: 7
    fig_width: 10
---
test
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
```

This code is developed to visualize all the spectra in an appropriate manner. For this purpose we use the dps1200 dataframe. 
```{r data_definition}

# Save the data as a new dataframe
data<-dps1200

head(data)

```

```{r clean-up}

# Correct column names by removing 'X' at the beginning
names(data) <- gsub("X", "", names(data))

head(data)
```
## Plotting of the spectral data

To visualize the data, different plots will be generated: 

* plot the absorbance against the wavelength

```{r plot-spectra}
library(ggplot2)

# save the spectral data as a new dataframe, without the descriptive coloumns
data_plot<-data[-(1:4)]

# Gathering the data 

# data_plot_long <- gather(data_plot, key = "wavelength", value = "intensity") # gather was reworked to pivot_longer
	# gather() is a function. reworked into pivot_longer()

data_plot_long <- pivot_longer(data_plot,cols = everything(),  names_to="Wavelength [nm]", values_to = "Intensity")

# change the datatype to numeric, for the wavelengths to be displayed correctly
data_plot_long$`Wavelength [nm]`<-(as.numeric(data_plot_long$`Wavelength [nm]`))

# for
head(data_plot_long)

# generate the plots for the spectras
ggplot(data_plot_long, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(alpha = 0.005, size = 1) + # , colour = 3) +
  labs(title = "Overview of the spectra",
       x = "Wavelength [nm]",
       y = "Intensity [1]")

# alpha = 0.005 defines the transparency, points overlapping can be better differentiated

ggplot(data_plot_long, aes(x = `Wavelength [nm]`, y = Intensity)) +
  geom_point(alpha = 0.005, size = 1) + # , colour = 3) +
  labs(title = "Overview of the spectra",
       x = "Wavelength [nm]",
       y = "Intensity [1]")

```

```{r}
ggplot(data, aes(x = tree)) +
 geom_line(aes(y = `2970`, color = "2970")) +
 geom_line(aes(y = `2883`, color = "2883")) +
 scale_color_manual(values = c("2970" = "blue", "2883" = "red")) +
 labs(x = "Year", y = "Value", color = "Column", title = "")+
	theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```


```{r}
ggplot(data, aes(x = type, fill=Origin)) + 
	geom_bar()

```


```{r}
# Generate a faceted bar plot by Origin
ggplot(data, aes(x = type)) +
 geom_bar() +
 facet_wrap(~Origin) +
 labs(x = "Type", y = "Count", title = "Count of Types by Country")

# Create the faceted bar plot by Type of Environment
ggplot(data, aes(x = Origin)) +
 geom_bar() +
 facet_wrap(~type) +
 labs(x = "Origin", y = "Count", title = "Count of Origins by Type of Environment")

# Create the single bar plot colored by country
ggplot(data, aes(x = type, fill = Origin)) +
 geom_bar(position ="dodge2") +
 labs(x = "Type", y = "Count", fill = "Country", title = "Count of Types by Country")

# Create the single bar plot colored by country
```


```{r}
# Find the position of the 'type' column
type_position <- which(names(data) == "type")
type_position

# Select all columns after the 'type' column
numerical_columns <- names(data)[type_position + 1:ncol(data)]

# Print the first five selected columns
head(numerical_columns)

# Initialize an empty ggplot object
p <- ggplot()

# Loop through each pair of numeric columns
# for (i in 1:(length(numerical_columns) - 1)) {
#  for (j in (i + 1):length(numerical_columns)) {
#    # Create a plot for each pair
#    plot <- ggplot(data, aes_string(x = numerical_columns[i], y = numerical_columns[j], color = "type")) +
#      geom_point() +
#      labs(x = numerical_columns[i], y = numerical_columns[j], color = "Type") +
#      theme_minimal()
#    
#    # Print the plot
#    print(plot)
# }
#}
```


```{r}
# loop over the unique values in data$type
# create the plot (x= numerical_columns, y = value), color by

# ggplot(data, aes(x=numerical_columns[1], y = value))

# Initialize an empty ggplot object
p <- ggplot()

for (i in unique(data$type)){
	# 
}

# Loop over each numerical column
for (col in names(data)[which(names(data) %in% numerical_columns)]) {
 # Add points to the plot for the current column
 p <- p + geom_point(data = data, aes(x = col, y = data[[col]], color = type))
}

# Add labels and title
p <- p + labs(x = "Wavelength", y = "Intensity", color = "Type", title = "NIR-Spectra Intensity by Wavelength and Type")

# Display the plot
print(p)

```
```{r}
# Load necessary libraries
library(ggplot2)
library(tidyverse)

# Assuming 'data' is your dataframe and 'numerical_columns' contains the names of your wavelength columns
# Reshape the data to a long format
data_long <- data%>%
 pivot_longer(cols = all_of(numerical_columns, names_to = "Wavelength", values_to = "Intensity"))

# Plot the reshaped data
p <- ggplot(data_plot, aes(x = data_plot$, y = "Intensity", color = "type")) +
 geom_point() +
 labs(x = "Wavelength", y = "Intensity", color = "Type", title = "NIR-Spectra Intensity by Wavelength and Type")

# Display the plot
print(p)

```


```{r coloured plot}
head(data)

ggplot(data, aes(x = year, y = `2970`)) +
 geom_point() +
 labs(x = "Year", y = "Value", title = "Scatter Plot of 2970 Over Years")



for(i in unique(data$Origin)){
	ggplot(data = data, aes(data$type, count(i)))
}
```


## Plotting of the age/year

for each tree we have multiple spectral measurements, where each measurement includes a specific tree ring. To get a sense for the age distribution, the Quartiles will be calculated. 

```{r}
ggplot(data, aes(x = year)) +
 geom_bar(fill = "steelblue") +
 theme_minimal() +
 labs(x = "Year", y = "Count", title = "Count of Measurements per Date of Wood growth")
```


```{r}
ggplot(data, aes(x = year)) +
 geom_density(fill = "steelblue") +
 theme_minimal() +
 labs(x = "Year", y = "Count", title = "Wood Year Distribution")
```


```{r}
ggplot(data, aes(x = Origin)) +
 geom_bar() +
	theme_minimal() +
 labs(x = "Origin", y = "Count", title = "Measurements per country")
```
