# extract the dataframe 

This code is developed to extract the data from the environment ('dataPSsub_new3.RData') to enable the option to work with the data in python. 

##  set up the Environment
```{r include=FALSE}

# clear previously loaded data
rm(list = ls())

knitr::opts_chunk$set(message = FALSE)
```
for clarification, we specify the directory structure, which is currently set up  

setwd("C:/Users/bernd/Dropbox/Masterarbeit_2023/1_Statistical_Learning_Spangl/data")

getwd()
[1] "C:/Users/bernd/Dropbox/Masterarbeit_2023/1_Statistical_Learning_Spangl/data"

```{r directory-structure}
home_dir<-"C:/Users/bernd/Dropbox/Masterarbeit_2023/1_Statistical_Learning_Spangl/"
code_dir<- "../code/"
data_dir<-"../data/extracted_from_rdata/"

```

## Specification of the data we are intrested in

the cleaned data, we are interested in are: 
'dps1200' 
'dpsDeriv1200' 
aus dem RData-File 'dataPSsub_new3.RData' 

## Manual Import
1. Open the Rdata-file
2. save the files as .csv

## automated Import

```{r}

load("C:/Users/bernd/Dropbox/Masterarbeit_2023/1_Statistical_Learning_Spangl/code/Code_Bernhard/dataPSsub_new3.RData")
```

The paste0() function concatenates strings without spaces to generate a valid file path.

row.names={FALSE or TRUE} argument is optional and specifies if the row names should be written to the CSV file

```{r write-data}
# saving the data from the dps1200 and dpsDeriv1200 dataframe as singular data
if(file.exists(data_dir)){
	print('FILE: dps1200.csv exists already')
} else {
	write.csv(dps1200, paste0(data_dir, "dps1200.csv"), row.names=TRUE)
	print('FILE: dps1200.csv writen')
}

if(file.exists(data_dir)){
	print('FILE: dpsDeriv1200 exists already')
} else {
	write.csv(dpsDeriv1200, paste0(data_dir, "dpsDeriv1200"), row.names=TRUE)
	print('FILE: dpsDeriv1200 writen')
}
# write.csv(dpsDeriv1200, paste0(data_dir, "dpsDeriv1200.csv"), row.names=TRUE)

# To keep only the environmental data "dps1200", we will combine the rm command with the setdiff() (difference between two sets, here the list generated with ls() and the list containing the data to keep)
```

Discard unused Data

```{r clean-up}
# Define the variables to keep in the working environment
keep<-c("home_dir", "data_dir", "code_dir", "dps1200",  "dpsDeriv1200")

# remove all variables, which are not in the list "keep" list
rm(list = setdiff(ls(), keep))

```
